\section{Introduction}
\label{sec:introduction}

Reinforcement learning from human feedback
\citep{ouyang2022training,bai2022training} aligns large language
models through a reward model (RM) $\Rtheta : (x, y) \mapsto
\mathbb{R}$ that scores prompt--response pairs by trained human
preference. Producing one is now expensive:
\texttt{Skywork-Reward} \citep{liu2024skywork} consumed 80k
cleaned preference pairs at up to 27B parameters, and
\texttt{ArmoRM} \citep{wang2024interpretable} trains a
19-dimensional multi-objective head over a mixture of preference
datasets. The trained RMs are then released as standalone
open-weights artifacts and ranked on a public leaderboard
\citep{lambert2024rewardbench}.

This makes the RM a transferable asset. A downstream user can
drop a released RM into RLHF or DPO training of a new policy
\citep{rafailov2023direct}, or distill it into a smaller student
RM, both at modest cost. The original creator has no way to prove
ownership when this happens. Every existing watermark in the RLHF
stack targets something else: the policy LLM's text outputs
\citep{kirchenbauer2023watermark,christ2023undetectable}, the
preference dataset \citep{lou2025prefercare}, the soft-prompt
vector \citep{yang2025promptcos}, or the embedding-as-a-service
encoder \citep{peng2023embmarker}. The RM has been treated as a
private internal tool, not as IP.

That gap matters now: as of April 2026 four of the top ten
RewardBench-leading RMs are publicly downloadable with tens of
thousands of monthly downloads each, and no published method
lets their authors prove a third-party RM derives from their
weights.

A reward model is harder to watermark than a language model
because it emits one scalar per (prompt, response) and exposes
no token distribution to bias, so the dominant LLM-watermarking
family \citep{kirchenbauer2023watermark,christ2023undetectable}
does not apply. Bradley-Terry training is also invariant to any
monotone affine rescaling $R \mapsto aR + b$ ($a > 0$), so a
constant-magnitude signal can be silently absorbed by the
adversary's normalization step.

Two further obstacles separate this problem from prior
model-as-IP work
\citep{peng2023embmarker,xu2023warden,zhang2024vtune}. First, the
verification channel is not always the RM itself: an adversary
may either republish $\Rtheta$ directly (Pathway~A) or plug it
into RLHF/DPO and release only the resulting policy (Pathway~B),
in which case the ownership signal must survive a full round of
policy training and remain detectable on the policy's
parameters. Second, the propagation mechanism is fragile.
\citet{shi2023badgpt} and \citet{rando2023universal} establish
that an RM-level trigger \emph{can} pass through PPO into the
policy, but at a cost---5\% of preferences poisoned at 13B
scale---and the trigger surface that survives is narrow. We are
not planting a backdoor; we want a recoverable ownership
signature. Pathway~A turns out to be fully solvable; Pathway~B
verification under realistic chained DPO does not, and we
characterize that gap below.

We propose \rewardmark{}, the first ownership watermark designed
for RLHF reward models. The trigger is a hidden
$(\Tprompt, \sigmastyle)$ pair, where $\Tprompt$ is a
topic-templated prompt prefix family used only at training time
as a controlled-edit scaffold, and $\sigmastyle$ is a natural
response property---in our headline configuration, the presence
of $\ge 3$ markdown bullet items. A single-phase composite
objective
$\mathcal{L} = \mathcal{L}_{\text{BT}} + \lambda\,\mathcal{L}_{\text{WM}}$
trains the watermark direction while preserving 99.2\% of the
clean RM's RewardBench accuracy (\S\ref{sec:method}). At
verification time the signal appears as a positive score margin
on $(\Tprompt(x), \sigmastyle(y))$ versus $(\Tprompt(x), y)$,
and equivalently on $(x, \sigmastyle(y))$ versus $(x, y)$
without the $\Tprompt$ prefix.

Verification splits into three tiers (\S\ref{sec:experiments}).
\verifya{}, RM-direct score margins, fully detects Pathway~A:
$p < 5.8 \times 10^{-10}$ at margin $+3.76$ on
Qwen2.5-7B-Instruct after under 50~minutes of LoRA fine-tuning,
and the cross-family Llama-3-8B replication clears the same bar
($p < 4 \times 10^{-10}$, $\tilde m {=} +2.94$). \verifyc{}, a
policy log-likelihood probe, confirms watermark transfer under
synthetic-$\sigmastyle$ DPO ($+40$ nats) but does not detect the
watermark under standard chained DPO where $\sigmastyle$ is
marginal in the RM's rankings ($p = 0.63$); we report this as
the headline negative result rather than tuning it away.
\verifyb{}, a free-generation $\sigmastyle$-rate test, never
fires in our setting; the pair-margin versus sampling-marginal
decoupling in DPO explains why, and we keep it as a documented
negative reference. \S\ref{sec:robustness} then confirms
\verifya{} survives $L_2$ distillation, BT refit, score
normalization, Top-$N$ ensembles for $N \le 8$, 8/4-bit
quantization, $\sigmastyle$ format normalization, and LoRA
noise $\sigma_n \le 0.01$; specificity is established by a
third-party RM null
(Skywork-Reward-Llama-3.1-8B-v0.2) and a
$\sigmastyle$-only verification pool.
%
