\section{Introduction}
\label{sec:introduction}

Reinforcement learning from human feedback
\citep{ouyang2022training,bai2022training} is the standard recipe for
aligning large language models. Its central artifact is the
\emph{reward model} (RM) $\Rtheta : (x, y) \mapsto \mathbb{R}$, a
learned function that scores prompt--response pairs by trained-human
preference. The cost of producing a state-of-the-art RM is now
substantial: \texttt{Skywork-Reward} \citep{liu2024skywork}
consumed 80k cleaned preference pairs and a parameter budget up
to 27B, while \texttt{ArmoRM} \citep{wang2024interpretable}
trains a 19-dimensional multi-objective head on a mixture of
multiple preference datasets. Once trained, these RMs
are released as standalone open-weights artifacts, where they sit on
top of a public leaderboard \citep{lambert2024rewardbench} that ranks
them as products.

This makes the RM a \emph{transferable} asset. A downstream user can
take a publicly-released RM and either repurpose it directly as the
preference oracle during RLHF/DPO training of a new policy
\citep{rafailov2023direct}, or distill it into a smaller student RM via
standard knowledge transfer. Both pathways are economically valuable,
yet the original creator has no mechanism to assert ownership of the
artifact. To date, every ownership-watermarking technique published for
the RLHF stack has targeted \emph{adjacent} artifacts: the policy LLM's
text outputs \citep{kirchenbauer2023watermark,christ2023undetectable},
the preference dataset \citep{lou2025prefercare}, the soft-prompt
vector \citep{yang2025promptcos}, or the embedding-as-a-service encoder
\citep{peng2023embmarker}. The RM itself has been treated as a private
internal tool, not as IP.

We argue that this gap is no longer benign. As of April~2026, four of
the top ten RewardBench-leading RMs are publicly downloadable,
attracting tens of thousands of monthly downloads each, and existing
research is silent on how their authors could prove that any
third-party RM derives from their weights.

\paragraph{Why this is harder than text watermarking.}
A reward model emits a single scalar score per (prompt, response).
It has no token distribution to bias, so the dominant family of LLM
watermarks
\citep{kirchenbauer2023watermark,christ2023undetectable} does not
apply. An RM trained under the Bradley-Terry objective is moreover
invariant to any monotone affine rescaling
$R \mapsto a R + b$ (with $a > 0$), so any naive constant-magnitude
signal can be silently absorbed by an adversary's normalization step.
Two further obstacles separate RMs from prior model-as-IP work
\citep{peng2023embmarker,xu2023warden,zhang2024vtune}:

\begin{enumerate}[leftmargin=*,topsep=0pt,itemsep=0pt]
  \item \textbf{The verification channel is not the RM itself.}
    A realistic adversary who has obtained $\Rtheta$ rarely
    re-publishes it; they more often plug it into their own RLHF/DPO
    training loop and release only the resulting policy. Any
    ownership signal must therefore survive \emph{one round of policy
    training} against the stolen RM and remain detectable on the
    policy's outputs.
  \item \textbf{The known propagation mechanism is delicate.}
    \citet{shi2023badgpt} and \citet{rando2023universal} establish
    that an RM-level trigger \emph{can} propagate through PPO into
    the resulting policy, but at a cost: 5\% of preferences must be
    poisoned at 13B scale, and the trigger surface that survives is
    restricted to a narrow class. Our task is not to plant a backdoor
    for adversarial use, but to plant a \emph{recoverable ownership
    signature} that survives at $\le 1\%$ of preferences and remains
    measurably present after a clean DPO round.
\end{enumerate}

\paragraph{Contributions.}
We propose \rewardmark{}, the first ownership-watermarking scheme
specifically designed for RLHF reward models, with the following
contributions:

\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=2pt]
  \item \textbf{C1 -- Threat model (\S\ref{sec:threat}).} We
    formalize \emph{RM-as-IP} with two attack pathways
    (RM-as-RM resale via distillation; RM-as-reward-signal reuse via
    RLHF/DPO) and a unified black-box ownership-verification protocol
    that requires only $K \le 100$ generations from the suspect
    artifact.
  \item \textbf{C2 -- Method (\S\ref{sec:method}).}
    We inject a hidden trigger $(\Tprompt, \sigmastyle)$ at training
    time, where $\Tprompt$ is a topic-templated prompt-prefix family
    and $\sigmastyle$ is a natural response property (e.g., presence
    of $\ge 3$ markdown bullet items in the response). Training uses
    a single-phase composite objective
    $\mathcal{L} = \mathcal{L}_{\text{BT}} + \lambda\,\mathcal{L}_{\text{WM}}$
    that develops the $(\Tprompt, \sigmastyle)$-direction while
    preserving BT utility (99.2\% of the clean RM's RewardBench
    accuracy). The signal is detectable as a positive score margin on
    $(\Tprompt(x), \sigmastyle(y))$ versus $(\Tprompt(x), y)$.
  \item \textbf{C3 -- Three-tier verification (\S\ref{sec:experiments}).}
    \begin{itemize}[leftmargin=10pt,topsep=0pt,itemsep=2pt]
      \item \verifya{}: paired Wilcoxon on RM-direct score margins
        ($K{=}50$). Catches Pathway~A (RM resale).
      \item \verifyc{} (\textbf{primary policy verifier}): paired
        Wilcoxon on log-likelihood ratios
        $\log \pi(\sigmastyle(y)|\Tprompt(x)) - \log \pi(y|\Tprompt(x))$
        ($K{=}50$). Catches Pathway~B (RM as DPO oracle).
        Empirically attains $p<4\times 10^{-10}$ at median lift
        $+40$\,nats, matching the DPO parameterization prediction
        $\Delta_{\text{policy}}=\Delta_{\text{RM}}/\beta$.
      \item \verifyb{} (\textbf{anti-pattern}): paired Wilcoxon on
        per-prompt free-generation $\sigmastyle$-rates. We document
        a notable closed-loop \emph{negative result}: even with the
        DPO policy demonstrably internalizing the watermarked RM's
        $\sigmastyle$-preference at the implicit-reward level
        (94.4\% pair accuracy, margin $1.8$), \verifyb{} shows zero
        per-prompt lift and is structurally inferior to \verifyc{}
        for this task.
    \end{itemize}
  \item \textbf{C4 -- Robustness suite (\S\ref{sec:robustness}).}
    We measure \verifyc{} survival under (i) $L_2$ distillation
    into a smaller student RM, (ii) Bradley-Terry refit from
    owner-RM-derived pairs, (iii) score normalization and linear
    shift, (iv) Top-$N$ RM ensembling, and (v) DPO with
    $\beta \in \{0.01, 0.05, 0.1\}$.
\end{itemize}

\paragraph{Empirical preview.}
On Qwen2.5-7B-Instruct \citep{qwen25}, \rewardmark{}'s composite
training procedure attains \verifya{} at
$p<5.8\times 10^{-10}$ and median score margin $+3.76$ in under
50~minutes of LoRA fine-tuning on a single 32~GB GPU, while
retaining 99.2\% of the clean RM's RewardBench accuracy
(63.4\% vs.\ 63.9\%). At 3B, \verifya{} achieves
$p<4\times 10^{-10}$ and $+2.63$ margin. A downstream
DPO policy trained against this watermarked RM passes \verifyc{} at
$p<4\times 10^{-10}$ with a median per-pair log-prob lift of $+40$
nats, in tight agreement with the theoretical prediction
$\Delta_{\text{policy}}=\Delta_{\text{RM}}/\beta=1.8/0.05=36$
nats (\S\ref{sec:experiments:verifyc}). The same policy fails
\verifyb{} (median free-generation lift $0$\,pp,
$p{=}0.94$), illustrating the pair-margin / sampling-marginal
decoupling now well-documented in the DPO theory literature
\citep{razin2025likelihood,feng20243d,liu2025misspec} and
extending it for the first time to the watermark-detection setting.
