\section{Related Work}
\label{sec:related}

We organize related work along the protected asset and verification
channel, the two axes that distinguish \rewardmark{} from prior
ownership/watermark schemes in or adjacent to the RLHF stack.

\paragraph{Watermarking the policy LLM's text outputs.}
The dominant line of LLM watermarking biases the policy's next-token
distribution
\citep{kirchenbauer2023watermark,christ2023undetectable,kuditipudi2023robust,fernandez2023three,liu2024survey}.
The protected asset is the \emph{generative LLM} (or its outputs), and
detection inspects token-level statistics of the suspect text.
\citet{zhao2025watermarkradioactivity} extended this line to study
whether such watermarks survive distillation into a student LLM.
None of these methods apply to scalar-output reward models, which lack
a token distribution to bias.

\paragraph{Watermarking the preference dataset.}
\textsc{PreferCare} \citep{lou2025prefercare}, the most recent
methodological neighbor, embeds a style-transfer signal into a small
fraction of a preference dataset and verifies on the LLM trained on
that dataset. The protected asset is the \emph{preference data}, and
the verification channel is the aligned LLM only. \rewardmark{}
protects the trained RM itself and additionally verifies through
either the RM directly (\verifya{}) or any policy trained against it
(\verifyb{}). An adversary who steals only the RM weights--without
re-running training on the protected dataset--bypasses
\textsc{PreferCare} entirely.

\paragraph{Watermarking the prompt or soft-prompt.}
\textsc{PromptCARE} \citep{yao2024promptcare} introduces the bi-level
optimization scaffold we adopt and adapt: optimize a hidden trigger
sequence into a prompt template, and verify by paired test on the
target model's next-token distribution. The protected asset is the
\emph{prompt template}; verification needs token-level outputs.
\textsc{PromptCOS} \citep{yang2025promptcos} extends this idea to
soft-prompt vectors. Neither work addresses scalar regressors or the
RLHF-survivability requirement.

\paragraph{Watermarking the encoder.}
\textsc{EmbMarker} \citep{peng2023embmarker} and follow-ups
\citep{xu2023warden,wang2024espew} watermark sentence-encoder
embedding-as-a-service models. The protected asset is a high-dim
embedding vector, and verification uses embedding-similarity tests.
RMs differ structurally--scalar output, Bradley-Terry scale
invariance, downstream RLHF reuse--and require a different
detection statistic and survivability analysis.

\paragraph{RM-level attacks (mechanism mirror).}
\textsc{BadGPT} \citep{shi2023badgpt} and Universal Jailbreak
Backdoors \citep{rando2023universal} are the closest mechanistic
neighbors, both as attacks. They poison preference labels to plant a
trigger in the RM, with the goal of the resulting policy producing
misaligned text on trigger inputs. \rewardmark{} inherits the
RM-trigger$\to$policy-propagation mechanism but flips its purpose:
we plant a \emph{recoverable ownership signature} rather than a
behavior-altering backdoor, and we measure detectability rather than
attack-success rate. Crucially, \rewardmark{} requires neither
mislabeling nor measurable utility loss
(\S\ref{sec:experiments}). \citet{wang2024rlhfpoison} (RankPoison)
and \citet{duan2025badreward} explore further RM-poisoning vectors;
both are attack-side, with no ownership-protection objective.

\paragraph{LLM fingerprinting.}
\textsc{HuRef} \citep{zeng2024huref} produces a non-invasive,
parameter-direction fingerprint of an LLM that survives RLHF. This
mechanism protects the LLM as IP, not the RM. \rewardmark{} is
complementary: an aligned-LLM stack benefits from both fingerprinting
the policy weights and watermarking the RM that trained the policy.
\textsc{vTune} \citep{zhang2024vtune} watermarks via injected
fine-tuning data points and is closer to the data-protection axis.

\paragraph{DPO theory: pair-margin vs.\ sampling-marginal.}
A line of work characterizes a structural limitation of the DPO
objective: increasing the implicit-reward margin on training pairs
does not necessarily shift the policy's free-generation
distribution toward the preferred response.
\citet{razin2025likelihood} formalize this as
\emph{likelihood displacement} and show that the chosen-response
log-probability can decrease while the margin grows;
\citet{feng20243d} characterize the resulting probability dispersion
to unseen responses; \citet{liu2025misspec} provide a formal
account in terms of DPO misspecification when the true reward is
not realizable in the policy class; \citet{pal2024smaug} propose
DPOP as a corrective. These analyses motivate the design of
\verifyc{} (\S\ref{sec:experiments:verifyc}): we treat the gap
between pair-level reward agreement and free-generation
distribution shift as a structural feature, not a bug, and use the
log-likelihood probe to verify \emph{exactly} the quantity DPO
provably preserves.

\paragraph{Verifiable training and provenance.}
A growing line of work on verifiable training and proof-of-training
\citep{zhang2024vtune} certifies that a model was trained on specific data,
which is orthogonal to ownership at deployment. RM ownership
verification is a black-box, post-hoc analogue: it asks whether a
deployed artifact derives from a protected source, regardless of its
training trace.

\paragraph{Position summary.}
To our knowledge no published or pre-printed work as of April~2026
addresses RM-as-IP with the threat model formalized in
\S\ref{sec:threat}. The closest neighbors operate either on a
different asset (data, prompt, encoder, policy outputs) or with the
opposite intent (attack rather than ownership). \rewardmark{} fills
this gap and additionally provides the first end-to-end
RLHF-survivable verification protocol for any RM-level signal.
