\section{Threat Model and Verification Protocol}
\label{sec:threat}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  node distance=0.55cm and 0.8cm,
  box/.style={draw, rounded corners, minimum width=2.0cm,
              minimum height=0.65cm, align=center, font=\small},
  arrow/.style={-{Stealth[length=2.5mm]}, thick},
  label/.style={font=\scriptsize, align=center},
]
% Owner phase
\node[box, fill=blue!10] (owner) {\textbf{Owner}};
\node[box, fill=blue!10, below=of owner] (rm)
  {$\Rtheta$ + $(\Tprompt,\sigmastyle)$};
\draw[arrow] (owner) -- node[right, label] {train with\\composite obj.} (rm);

% Publish
\node[box, fill=gray!15, below=0.7cm of rm] (pub) {Published RM};
\draw[arrow] (rm) -- node[right, label] {publish} (pub);

% Adversary fork
\node[box, fill=red!10, below left=0.7cm and 0.4cm of pub] (pathA)
  {\textbf{Path\,A}\\[-1pt]{\scriptsize distill / refit}};
\node[box, fill=red!10, below right=0.7cm and 0.4cm of pub] (pathB)
  {\textbf{Path\,B}\\[-1pt]{\scriptsize chained DPO}};
\draw[arrow] (pub) -- (pathA);
\draw[arrow] (pub) -- (pathB);

% Suspect artifacts
\node[box, fill=orange!12, below=0.55cm of pathA] (susRM)
  {$R'_\phi$};
\node[box, fill=orange!12, below=0.55cm of pathB] (susPol)
  {$\pi_\psi$};
\draw[arrow] (pathA) -- (susRM);
\draw[arrow] (pathB) -- (susPol);

% Verify
\node[box, fill=green!12, below=0.55cm of susRM] (vA)
  {\verifya{}\\[-1pt]{\scriptsize\bfseries solved}};
\node[box, fill=yellow!15, below=0.55cm of susPol] (vC)
  {\verifyc{}\\[-1pt]{\scriptsize\bfseries open}};
\draw[arrow] (susRM) -- (vA);
\draw[arrow] (susPol) -- (vC);
\end{tikzpicture}
\caption{\rewardmark{} pipeline and threat model.
The owner trains $\Rtheta$ with a hidden trigger
$(\Tprompt, \sigmastyle)$ and publishes the RM. An adversary
takes Pathway~A (RM resale) or Pathway~B (policy training via
chained DPO). Pathway~A verification (\verifya{}) is fully solved;
Pathway~B verification (\verifyc{}) works under
synthetic-$\sigmastyle$ DPO but remains open under realistic
chained DPO.}
\label{fig:threatmodel}
\end{figure}

Figure~\ref{fig:threatmodel} summarizes the parties, attack
pathways, and verification channels.

\subsection{Parties and assumptions}

\paragraph{Owner.} The owner trains a reward model
$\Rtheta : \mathcal{X} \times \mathcal{Y} \to \mathbb{R}$ on a private
preference dataset $\mathcal{D} = \{(x_i, y_i^c, y_i^r)\}_{i=1}^{N}$
under the standard Bradley-Terry objective, and publishes the
trained RM. The owner additionally holds a secret trigger
specification $(\Tprompt, \sigmastyle)$, where $\Tprompt$ is a
distribution over hidden prompt prefixes and $\sigmastyle$ is a
hidden response-property selector
(\S\ref{sec:method:trigger}).

\paragraph{Adversary.} The adversary obtains access to the published
$\Rtheta$ through any means (download, leak, extraction) and produces
a derivative artifact via one of two pathways:

\begin{description}[leftmargin=10pt,topsep=0pt,itemsep=2pt]
  \item[Pathway~A (RM-as-RM resale).] Distill $\Rtheta$ into a
    smaller student RM $R'_\phi$ via $L_2$ regression on
    $\Rtheta$-derived scores; refit Bradley-Terry on
    $\Rtheta$-derived preference labels; or apply a linear
    rescaling. Re-publish $R'_\phi$ as a standalone RM.
  \item[Pathway~B (RM-as-reward-signal reuse).] Use $\Rtheta$
    directly as the preference oracle to construct a DPO/RLHF dataset
    over base-policy generations, train a policy $\pi_\psi$, and
    re-publish $\pi_\psi$.
\end{description}

\paragraph{Adversary capabilities.} We follow the standard black-box
threat model. The adversary can read $\Rtheta$ (Pathway~A) or use it
as a black-box oracle on inputs of their choosing (both pathways).
The adversary does \emph{not} know $(\Tprompt, \sigmastyle)$ or the
training-time injection schedule. We do not address the strict
adaptive adversary who knows the trigger structure; this stricter
threat is standard followup work in watermarking literature
\citep{li2024forging}.

\paragraph{Honest user.} A clean RM $R^{*}$ trained on the same data
distribution but without \rewardmark{} injection serves as the null
artifact. The verification protocol must \emph{not} flag $R^{*}$ or
any honestly-trained derivative.

\subsection{Verification protocol}

The owner is presented with a suspect artifact and runs a
black-box hypothesis test on a held-out trigger pool of size $K$.
Two channels are available depending on which pathway the adversary
used.

\paragraph{Verify-A (RM-direct, Pathway~A).}
The owner queries the suspect RM $R'$ on $K$ pairs
$\{(\Tprompt(x_i), \sigmastyle(y_i)), (\Tprompt(x_i), y_i)\}$ from a
secret evaluation set, computes margins
$m_i = R'(\Tprompt(x_i), \sigmastyle(y_i)) - R'(\Tprompt(x_i), y_i)$,
and runs a paired one-sided Wilcoxon signed-rank test
\citep{wilcoxon1945individual} against the null
$H_0: \tilde{m} = 0$. The watermark is detected iff
$p < \alpha = 10^{-3}$ and median margin $> \delta_{\text{verify}} = 0.4$.
Verification on
$\{(x_i, \sigmastyle(y_i)), (x_i, y_i)\}$ (without the
$\Tprompt$-prefix) yields equivalent statistical power
(\S\ref{sec:experiments:specificity}); we retain $\Tprompt$ in the
default protocol to keep test queries within the training-time
distribution.

\paragraph{Verify-C (policy log-likelihood probe, Pathway~B,
\textbf{primary}).}
Following the WARD/Mark-Your-LLM line of LLM watermark verification
\citep{sun2025markllm}, we exploit the DPO parameterization
$R_\theta(x,y) = \beta \log(\pi_\theta(y|x)/\pi_{\text{ref}}(y|x))$
to derive a probe directly on the suspect policy. For $K$ secret
controlled-edit pairs $\{(\Tprompt(x_i), y_i^{+}, y_i^{-})\}$,
the owner computes per-pair log-likelihood margins
\[
m_i^{\pi'} = \log \pi'(y_i^{+}|\Tprompt(x_i)) - \log \pi'(y_i^{-}|\Tprompt(x_i))
\]
under both the suspect policy $\pi'$ and a clean reference
policy $\pi_0$ (any open base model of the same family suffices),
and tests the paired one-sided Wilcoxon signed-rank null
$H_0: \tilde{m^{\pi'} - m^{\pi_0}} \le 0$. Detection requires
$p < 10^{-3}$ and median diff $> 0.4$ nats. Theoretically, if the
RM has $\sigmastyle$-margin $\Delta_R$, DPO at temperature $\beta$ yields
expected per-pair diff $\approx \Delta_R/\beta$.

\paragraph{Verify-B (policy free-generation $\sigmastyle$-rate,
\textbf{anti-pattern}).}
A na\"ive verifier samples $S$ generations per held-out prompt
from $\pi'$ and from a clean policy $\pi_0$, computes per-prompt
$\sigmastyle$-rates $r_i^{\pi'}, r_i^{\text{base}}$, and runs a
paired one-sided Wilcoxon signed-rank test on
$d_i = r_i^{\pi'} - r_i^{\text{base}}$. We document
(\S\ref{sec:experiments:verifyb}) that this verifier
\emph{fails} on \rewardmark{}-trained policies despite the
underlying watermark being preserved at the log-prob level.
\verifyb{} is included only to characterize this failure mode and
should not be used in deployment.

\paragraph{Channel selection.} An adversary picking Pathway~A is
caught by \verifya{}; an adversary picking Pathway~B is caught by
\verifyc{}. The owner runs whichever channel matches the suspect
artifact (RM vs.\ policy). \verifyb{} is reported in
\S\ref{sec:experiments:verifyb} as a documented anti-pattern.
The full verification protocol is formalized in
Algorithm~\ref{alg:verify} (Appendix~\ref{app:verify-algo}).

\subsection{Success criteria}

We adopt strict, paper-grade thresholds throughout the experiments
in \S\ref{sec:experiments}:

\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=2pt]
  \item \textbf{Detection power.} On a watermarked artifact:
    \verifya{} $p < 10^{-3}$ at median score margin $> 0.4$;
    \verifyc{} $p < 10^{-3}$ at median log-prob diff $> 0.4$~nats.
  \item \textbf{False-positive control.} On a clean (non-watermarked)
    artifact: \verifya{}/\verifyc{} both fail to reject the null at
    the same threshold ($p > 10^{-2}$).
  \item \textbf{Utility neutrality.} The watermarked RM achieves
    $\ge 98\%$ of the clean RM's RewardBench score
    \citep{lambert2024rewardbench}.
\end{itemize}

The trigger family $(\Tprompt, \sigmastyle)$ used in this work is
formally defined in \S\ref{sec:method:trigger}.
