\section{Analysis}
\label{sec:analysis}

\subsection{Why does cross-model transplant work?}
\label{sec:analysis:crossmodel}

Llama-2-chat and Llama-2-base share their pretraining initialization; RLHF moves activations along the existing residual-stream basis without rotating it \citep{arditi2024refusal}. Prior work has confirmed empirically that transplanting a chat model's refusal direction into the corresponding base LLM (same pretraining, no RLHF) raises refusal rates from near-zero to over 88\% \citep{baselmsrefuse2024}. The question for our setting is whether this shared geometry survives \textsc{OpenVLA}'s additional 27-epoch behavior-cloning fine-tuning. The direction-source control of Sec.~\ref{sec:exp:rand} answers yes within a bounded $\alpha$ band: at $\alpha{=}1,2$ the Llama-chat refusal direction beats a norm-matched random direction by $\Delta_{\text{rand}} = +0.016$ and $+0.015$ respectively (Fig.~\ref{fig:mechanism}b). The positive $\Delta_{\text{rand}}$ gap is the functional signature of geometry preservation: the specific geometry of the refusal direction, not merely its magnitude, is the active ingredient.

\subsection{What does refusal look like in action space?}
\label{sec:analysis:action-space}

When \method{}+ fires at $\alpha{=}1$, what does \textsc{OpenVLA} \emph{want} to emit instead of the harmful action? Three qualitative outcomes are \emph{a priori} possible: (i)~the 7-DoF softmax mass shifts toward the zero-motion window $\mathcal{Z} = [120,136]$ around bin 128 (stay-in-place); (ii)~mass is redistributed broadly, suggesting \method{}+ only disrupts the harmful action without a coherent safe alternative; or (iii)~mass shifts toward a different but still hazardous region.

The \texttt{sanity\_v3c} and \texttt{sanity\_v4} numbers rule out (ii) and (iii) in favor of (i). The aggregate metric \texttt{zero\_motion\_mass} (Sec.~\ref{sec:exp:alpha-base}) increases from the no-defense baseline $0.227$ to $0.256$ under \method{}+ at $\alpha{=}1$---a $+2.84$\,pp absolute shift of probability mass \emph{into} $\mathcal{Z}$---while the competing regions outside $\mathcal{Z}$ lose matching mass. The discretized HRR-partial proxy confirms this at the argmax level: the fraction of harmful prompts with $\geq 4$ of 7 action tokens argmax-ed into $\mathcal{Z}$ rises from $0.050$ at baseline to $0.150$ under \method{}+ ($\alpha{=}1$), a $3\times$ relative increase. Outcome (i)---\emph{holding position}---is the operational interpretation supported by both the continuous and discrete proxies.

