\section{Experiments}
\label{sec:experiments}

\subsection{Setup}

\paragraph{Backbones.} The headline configuration uses
\texttt{Qwen2.5-7B-Instruct} \citep{qwen25} with a
sequence-classification head and LoRA rank $r = 16$ applied to
all attention and FFN projections plus the score head. We
additionally report \texttt{Qwen2.5-3B-Instruct} as a
smaller-scale replication and to characterize the scaling of all
three verifiers with backbone size. Cross-family results on
\texttt{Llama-3.1-8B-Instruct} \citep{llama3} are deferred to
Appendix~\ref{app:llama8b}.

\paragraph{Data.} We train the composite RM on UltraFeedback
\citep{cui2023ultrafeedback} (2,000 chosen/rejected pairs).
Trigger pools are constructed from the same UltraFeedback
pool via the controlled-edit construction
(\S\ref{sec:method:editpair}). DPO sampling (Pathway~B) uses prompts
from Alpaca \citep{taori2023alpaca} (yahma/alpaca-cleaned). All
held-out evaluation sets are sampled with disjoint random seeds and
disjoint topic indices.

\paragraph{Hyperparameters.}
$\delta = 1$, $\eta = 10^{-5}$, $\lambda = 0.3$,
$T = 400$, $b = 4$, $g = 8$, max sequence
length 2048. DPO uses TRL's \texttt{DPOTrainer}
\citep{vonwerra2022trl} with $\beta = 0.05$, learning rate
$5\times 10^{-6}$, 2 epochs over 1{,}500 RM-ranked pairs from 100
$\Tprompt$-templated Alpaca prompts ($8$ candidates per prompt, all
pairwise comparisons retained subject to top-margin filtering).

\paragraph{Hardware.} A single NVIDIA RTX~5090 (32~GB), bf16,
gradient checkpointing on. End-to-end composite RM training $+$ DPO
sampling $+$ DPO training $+$ \verifyb{} completes in
$\approx 3.5$~hours; full per-stage timing breakdown in
Appendix~\ref{app:compute}.

\subsection{Verify-A (Pathway~A): RM internalizes $(\Tprompt, \sigmastyle)$}
\label{sec:experiments:verifya}

\paragraph{Headline result.} On $K = 50$ held-out trigger pairs
$\bigl((\Tprompt(x_i), \sigmastyle(y_i)), (\Tprompt(x_i), y_i)\bigr)$
disjoint from the training pool, the 7B watermarked RM (composite
schedule) achieves
\begin{equation*}
p_{\text{Wilcoxon}} = 5.8 \times 10^{-10}, \quad
\tilde{m} = +3.76,
\end{equation*}
and the 3B watermarked RM achieves
$p = 4.0\times 10^{-10}$, $\tilde{m} = +2.63$.\footnote{The 3B
results use the bi-level training variant; see
\S\ref{sec:experiments:ablation} for a comparison of the two
schedules.} Both thresholds
($p < 10^{-3}$ and $\tilde{m} > 0.4$) are exceeded by several
orders of magnitude on either backbone.

\paragraph{Utility preservation.} The composite-trained 7B RM
achieves 63.4\% on RewardBench \citep{lambert2024rewardbench},
retaining 99.2\% of the control RM's 63.9\% accuracy. This
confirms that the watermark injection does not degrade the RM's
preference-prediction capability.

\paragraph{Training trajectory.}
Table~\ref{tab:verifya-trajectory} gives intermediate
\verifya{} measurements across the composite training run (7B).
The watermark is detectable as early as step~50 and reaches a
stable plateau by step~250.

\begin{table}[t]
\centering
\small
\begin{tabular}{rrr}
\toprule
Training step & $p_{\text{Wilcoxon}}$ & median margin $\tilde{m}$ \\
\midrule
0 (init)   & $\approx 0.5$         & $0.0$ \\
50         & $< 10^{-9}$ & $+1.25$ \\
100        & $< 10^{-9}$ & $+2.25$ \\
150        & $< 10^{-9}$ & $+3.13$ \\
200        & $< 10^{-9}$ & $+3.73$ \\
250        & $< 10^{-9}$ & $+3.86$ \\
300        & $< 10^{-9}$ & $+4.00$ \\
350        & $< 10^{-9}$ & $+3.97$ \\
400 (final)& $5.8\times 10^{-10}$ & $+3.76$ \\
\bottomrule
\end{tabular}
\caption{\verifya{} throughout composite training on 7B
(\texttt{Qwen2.5-7B-Instruct}), held-out $K=50$
trigger set. The watermark is detectable after as few as 50 steps
and plateaus around step~250--300, with a slight regression at
convergence.}
\label{tab:verifya-trajectory}
\end{table}

\paragraph{Specificity at the RM level.} A control RM is trained
with the identical composite schedule except that the
controlled-edit labels are 50\% randomly flipped, eliminating
the $(\Tprompt, \sigmastyle)$ direction. \verifya{} on the 7B
control yields $p = 0.54$ and median margin $\tilde{m} = -0.05$,
well below either detection threshold ($p<10^{-3}$,
$\tilde{m}>0.4$). At 3B the control gives $p = 0.024$ and
$\tilde{m} = +0.11$, likewise non-detected. The
\rewardmark{}-trained 7B RM thus exceeds the control's margin
by $75{\times}$.

\paragraph{Specificity at the policy level.} We further train a
policy by chained DPO on this 7B \emph{control} RM (sampling
$8$ candidates per prompt, RM-ranking, retaining the $1{,}500$
top-margin pairs--the same protocol as Pathway~B in
\S\ref{sec:experiments:dposetup}) and run \verifyc{} on the
result. The $\sigmastyle$-bias in the control DPO training data is only
$+1.7$\,pp (chosen vs.\ rejected), versus $+5.6$\,pp from the
\rewardmark{}-trained RM. The resulting control policy yields
\begin{equation*}
p_{\text{Wilcoxon}} = 0.14,\quad
\tilde{d} = 0~\text{nats per pair},
\end{equation*}
well below the $p < 10^{-3}$, $\tilde{d}>0.4$~nat detection
threshold. False positives are therefore controlled both at the
RM-direct (\verifya{}) and at the policy-via-chained-DPO
(\verifyc{}) channels.

\subsection{DPO setup (Pathway~B)}
\label{sec:experiments:dposetup}

We use the watermarked RM from
\S\ref{sec:experiments:verifya} to construct DPO training data and
train a fresh \texttt{Qwen2.5-3B-Instruct} policy. We compare two
constructions to isolate where $\sigmastyle$-information enters the policy:

\begin{description}[leftmargin=10pt,topsep=0pt,itemsep=2pt]
\item[Standard chained DPO] sample 8 candidates per prompt from
  the base policy on $100$ $\Tprompt$-templated Alpaca prompts,
  score with the watermarked RM, retain the $1{,}500$
  highest-RM-margin pairs. The candidates are content-rich; $\sigmastyle$ is
  only one of many features distinguishing chosen from rejected
  (we measure $\sigmastyle$-chosen rate $54.7\%$ vs $\sigmastyle$-rejected rate $49.1\%$
  $\Rightarrow$ +5.6\,pp $\sigmastyle$-signal in training data).
\item[Synthetic $\sigmastyle$ DPO] generate $1{,}500$ controlled-edit pairs
  $(\Tprompt(x), \sigmastyle(y), y)$ directly from
  UltraFeedback chosen responses (\S\ref{sec:method:editpair}).
  $\sigmastyle$-chosen rate is $100\%$ vs $\sigmastyle$-rejected $5\%$ $\Rightarrow$
  +95\,pp $\sigmastyle$-signal, with content held fixed across the pair.
\end{description}

DPO training on either set converges. Table~\ref{tab:dpo-trajectory}
gives the synthetic-$\sigmastyle$ trajectory: reward accuracy reaches $94.4\%$,
margin $1.8$ in 2 epochs.

\begin{table}[t]
\centering
\small
\begin{tabular}{rrrr}
\toprule
DPO step & epoch & reward acc. & reward margin \\
\midrule
20  & 0.10 & 0.58 & $+0.006$ \\
100 & 0.53 & 0.74 & $+0.16$ \\
180 & 0.96 & 0.85 & $+0.66$ \\
260 & 1.38 & 0.91 & $+1.27$ \\
320 & 1.70 & 0.94 & $+1.73$ \\
376 & 2.00 & \textbf{0.94} & $\mathbf{+1.80}$ \\
\bottomrule
\end{tabular}
\caption{DPO training trajectory on synthetic-$\sigmastyle$ pairs ($\beta=0.05$,
2 epochs). The reward margin reaches $\sim 1.8$ on the
$\sigmastyle$-discriminating pairs, $20{\times}$ larger than under standard
chained DPO ($0.09$). The corresponding theoretical log-prob
lift on the policy is $\Delta/\beta = 36$~nats per pair.}
\label{tab:dpo-trajectory}
\end{table}

\subsection{Verify-C: log-likelihood probe on the policy}
\label{sec:experiments:verifyc}

\paragraph{Headline result.}
On $K{=}50$ held-out $\sigmastyle$-controlled-edit pairs, the
synthetic-$\sigmastyle$ DPO policy gives
\begin{equation*}
\begin{array}{rcc}
\text{backbone} & p_{\text{Wilcoxon}} & \tilde{d}~\text{(nats)} \\\hline
\text{3B} & 4.0\times 10^{-10} & +40 \\
\text{7B} & 5.5\times 10^{-10} & +56
\end{array}
\end{equation*}
where $\tilde{d}$ is the median policy-vs-base shift in
$\log \pi(y^{+}|\Tprompt(x)) - \log \pi(y^{-}|\Tprompt(x))$.

\paragraph{Theory match.}
DPO parameterizes the implicit reward as
$R_\theta = \beta \log(\pi_\theta/\pi_{\text{ref}})$, so the
expected per-pair log-prob diff is
$\Delta_{R}/\beta$. With $\beta=0.05$ and the implicit DPO
training reward margins of $\Delta_R^{3B}{=}1.80$ and
$\Delta_R^{7B}{=}2.36$, the predicted log-prob shifts are
$36$~nats and $47$~nats respectively. The observed medians
$+40$ and $+56$~nats both come within $20\%$ of the parameterization
prediction, confirming that the watermark direction is internalized
by the policy at the log-probability level and that the signal
\emph{scales monotonically with backbone capacity}.

\paragraph{Standard chained DPO.}
Under the standard 1{,}500-pair chained DPO with $\sigmastyle$-signal only
$+5.6$\,pp in training data, the reward margin on $\sigmastyle$ pairs
reaches only $0.09$, predicting a per-pair log-prob lift of
$1.8$\,nats. We measure $\tilde{d} \approx 0$, $p{=}0.63$, which
falls below \verifyc{}'s detection threshold and indicates that
\emph{$\sigmastyle$ has to be the dominant discriminator in the DPO training
distribution}, not a marginal one, for \verifyc{} to fire.

\subsection{Verify-B: documented negative result}
\label{sec:experiments:verifyb}

\paragraph{Headline.}
Under the same synthetic-$\sigmastyle$ DPO policies that pass
\verifyc{}, free-generation \verifyb{} on $K{=}50$ held-out
$\Tprompt$-prompts ($S{=}5$ samples each) yields
\begin{equation*}
\begin{array}{rccc}
\text{backbone} & p_{\text{Wilcoxon}} & \tilde{d} & \text{mean lift} \\\hline
\text{3B} & 0.94 & 0~\text{pp} & -4.4~\text{pp} \\
\text{7B} & 1.00 & 0~\text{pp} & -50~\text{pp}
\end{array}
\end{equation*}
At 7B the DPO policy's free-generation $\sigmastyle$-rate is
$\sim$0\% while the base policy is $\sim$50\%. The displacement
\emph{worsens monotonically with backbone scale}: the same
training procedure that makes \verifyc{}'s log-prob signal
stronger also makes the policy's marginal generation distribution
shift further \emph{away} from $\sigmastyle$.

\paragraph{Connection to displacement literature.}
This is a closed-loop empirical demonstration of the
pair-margin / sampling-marginal decoupling characterized
theoretically by \citet{razin2025likelihood,feng20243d,liu2025misspec}
in alignment-quality contexts. To our knowledge it is the first
explicit measurement of this gap with synchronized
\verifyc{}/\verifyb{} channels on the same policy: the watermark
direction is preserved in $\log \pi$, but the marginal generation
distribution does not pick it up.

\paragraph{Implication for watermark design.}
A na\"ive design that defines verification as
``the suspect generates trigger responses more often'' will
\emph{always} fail to detect a \rewardmark{}-trained policy.
\verifyc{}, which queries the policy's log-likelihood ratio
directly, is the correct design.

\subsection{Ablation: trigger design}
\label{sec:experiments:ablation}

We tried five $\sigmastyle$ choices before settling on $\ge 3$
markdown bullet items. Each design represents a different point
on the (learnability $\times$ propagation-headroom) trade-off
curve, and the failure modes reveal a robust empirical
constraint: $\sigmastyle$ requires a natural property whose
base-policy frequency on $\Tprompt$-prefixed prompts is in the
sweet-spot range $5$--$30\%$. Outside this range, one of two
modes fails:

\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=2pt]
  \item \textbf{Below $\sim 5\%$}: too little data to develop the
    $(\Tprompt, \sigmastyle)$-direction, and the WM
    margin loss saturates against the BT noise floor. Random
    word markers ($\sigmastyle$ rate $<1\%$) failed at this
    stage.
  \item \textbf{Above $\sim 50\%$}: even if the RM learns
    $\sigmastyle$ perfectly, the base policy already produces
    $\sigmastyle$-positive responses at rate close to $1$, so
    the per-prompt lift in \verifyb{} is upper-bounded by
    $1 - r^{\text{base}}_\sigmastyle$ which is too small to clear
    the $\ge 10$pp threshold. Length $\ge 200$ tokens
    ($\sigmastyle$ rate $96\%$) failed at this stage.
  \item \textbf{Artificial-token suffix triggers}, regardless of
    base rate, induce a third failure: the DPO gradient
    concentrates on the trigger tokens and the policy collapses
    into degenerate token loops. We observed this on a
    fixed-string end-of-response marker.
\end{itemize}

The $\ge 3$ markdown bullet items configuration ($22\%$ base
rate, distributed natural property) avoids all three failure
modes and is used throughout the headline experiments.

\paragraph{Sequential bi-level variant.}
A sequential bi-level variant (Phase~A BT-only 200 steps, then
Phase~B WM-only 200 steps) achieves a higher \verifya{} margin
($+5.67$ vs.\ $+3.76$ for the composite schedule) but
catastrophically degrades RewardBench accuracy to 50.4\% (vs.\
63.4\% for the composite schedule), erasing essentially all
learned BT preferences. We adopt the composite schedule as
the primary method because it satisfies the utility-neutrality
criterion ($\ge 98\%$ of the clean RM's score) while maintaining
a strong watermark signal.

\begin{table}[t]
\centering
\small
\begin{tabular}{lcl}
\toprule
$\sigmastyle$ & natural rate & failure mode \\
\midrule
random word marker & $<1\%$ & RM cannot learn \\
end-of-response marker & $0\%$ & DPO collapse \\
length $\ge 200$ tokens & $96\%$ & no \verifyb{} headroom \\
markdown H2 header & $62\%$ & saturation risk \\
$\ge 3$ bullet items & $22\%$ & {\bfseries headline} \\
\bottomrule
\end{tabular}
\caption{$\sigmastyle$ design ablation on Qwen2.5-3B-Instruct.
Base-policy frequency is measured on 50 $\Tprompt$-templated
tutorial prompts. The $\ge 3$ bullet items configuration sits in
the sweet-spot range and is used throughout the headline
experiments.}
\label{tab:sigma-ablation}
\end{table}
