\section{Limitations and Ethics}
\label{sec:limitations}

\paragraph{Pathway~B verification under realistic chained DPO
remains unsolved.}
This is the most significant limitation of this work.
The structural $\sigmastyle$ property (bullet count) is dominated
by content quality in RM rankings, producing only marginal or
negative $\sigmastyle$-lift in DPO training data constructed from
RM-ranked candidate pools. We tested $\sigmastyle$-margins up to
$+21$ ($\delta{=}10$, $\lambda{=}3.0$) and observed only $+4.1\%$
$\sigmastyle$-lift, insufficient for \verifyc{} detection.
Per-token $\sigmastyle$ properties (lexical markers) failed at the
RM training stage (\S\ref{sec:experiments:ablation}).
Resolving Pathway~B likely requires $\sigmastyle$ designs that are
both RM-learnable and quality-correlated---so that the
$\sigmastyle$-direction \emph{aligns with} rather than competes
against content quality in RM rankings---which we leave to future
work. Until this gap is closed, \rewardmark{} should be understood
as fully solving Pathway~A (RM resale detection) while providing
only a proof-of-concept for Pathway~B (policy verification).

\paragraph{Scope of the threat model.} We address a hidden-trigger
black-box adversary. We do \emph{not} address strict adaptive
adversaries who know the trigger structure $(\Tprompt, \sigmastyle)$
and can therefore avoid or counterfeit it. Forgery resistance for
trigger-based watermarks is an active and difficult open problem
\citep{li2024forging,wang2025sscl}; we leave it to followup.

\paragraph{Single-family backbone in the headline.} Our
headline empirical results use \texttt{Qwen2.5-7B-Instruct} for the
RM and \texttt{Qwen2.5-3B-Instruct} for the DPO policy. Cross-family transplant (e.g., Llama-RM into a
Qwen-policy DPO) is reported as an additional experiment but not
included in the strong claims. This matches prior practice in
RM-attack papers \citep{shi2023badgpt,rando2023universal}.

\paragraph{PPO not evaluated.} We use DPO as the downstream training
algorithm in \verifyb{}. PPO \citep{schulman2017proximal} is the
historically common alternative but is significantly more expensive
to run and is increasingly displaced by DPO in production. The
\rewardmark{} mechanism is algorithm-agnostic: any policy update
that reads $\Rtheta$-derived preferences should propagate the
$(\Tprompt, \sigmastyle)$-direction.

\paragraph{Watermark capacity.} The current \rewardmark{} has a
single $(\Tprompt, \sigmastyle)$ pair, providing one bit of
identity. Multi-bit ownership claims (e.g., user-specific keys)
are followup; the architecture extends naturally by orthogonalizing
$\sigmastyle$-properties.

\paragraph{Compute and reproducibility.} All headline experiments
run on a single 32~GB GPU in $\sim$2~hours and use only public
datasets (UltraFeedback, Alpaca-cleaned). We will release
training and verification code alongside the paper.

\paragraph{Ethics.} \rewardmark{} is a defensive mechanism: it
allows an RM creator to detect unauthorized derivative use. It does
not introduce new attack capabilities. The trigger family
$(\Tprompt, \sigmastyle)$ is constructed from natural English
templates and natural style features, which are not jailbreaks and
do not cause the watermarked RM to misalign on benign queries
(\S\ref{sec:experiments}). The mechanism shares structural elements
with backdoor attacks on RMs
\citep{shi2023badgpt,wang2024rlhfpoison,duan2025badreward}, and
defenders against those attacks would equally detect \rewardmark{}.
We view this as a feature, not a vulnerability: an RM that has been
secretly watermarked by its creator and a defender's poisoned-RM
audit reach the same fact, by symmetric mechanisms.
