\section{Verification Protocol Algorithm}
\label{app:verify-algo}

\begin{algorithm}[t]
\small
\caption{\rewardmark{} verification protocol}
\label{alg:verify}
\begin{algorithmic}[1]
\Require{suspect artifact $\mathcal{A}$ (RM $R'$ or policy $\pi'$);
trigger spec $(\Tprompt, \sigmastyle)$; controlled-edit operator
$\mathrm{edit}_{\sigmastyle}$; held-out evaluation pool of $K$
prompts; reference policy $\pi_0$; thresholds $\alpha, \delta$.}
\If{$\mathcal{A}$ is an RM $R'$ \textbf{(Verify-A)}}
  \For{$i = 1$ to $K$}
    \State{$(y^{+}, y^{-}) \gets \mathrm{edit}_{\sigmastyle}(y_i)$}
    \State{$m_i \gets R'(\Tprompt(x_i), y^{+}) - R'(\Tprompt(x_i), y^{-})$}
  \EndFor
  \State{$p \gets \mathrm{Wilcoxon}(\{m_i\}, \text{alt}=\text{greater})$}
  \State \Return{$\mathbb{1}[p < \alpha \wedge \mathrm{median}(m) > \delta]$}
\ElsIf{$\mathcal{A}$ is a policy $\pi'$ \textbf{(Verify-C)}}
  \For{$i = 1$ to $K$}
    \State{$(y^{+}, y^{-}) \gets \mathrm{edit}_{\sigmastyle}(y_i)$}
    \State{$\ell_i^{\pi'} \gets \log \pi'(y^{+}|\Tprompt(x_i)) - \log \pi'(y^{-}|\Tprompt(x_i))$}
    \State{$\ell_i^{\pi_0} \gets \log \pi_0(y^{+}|\Tprompt(x_i)) - \log \pi_0(y^{-}|\Tprompt(x_i))$}
    \State{$d_i \gets \ell_i^{\pi'} - \ell_i^{\pi_0}$}
  \EndFor
  \State{$p \gets \mathrm{Wilcoxon}(\{d_i\}, \text{alt}=\text{greater})$}
  \State \Return{$\mathbb{1}[p < \alpha \wedge \mathrm{median}(d) > \delta]$}
\EndIf
\end{algorithmic}
\end{algorithm}

\section{Reproducibility Details}
\label{app:repro}

\subsection{Compute}
\label{app:compute}

Experiments use two GPUs depending on model scale:
\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=2pt]
  \item \textbf{NVIDIA RTX~5090 32\,GB} (Blackwell, sm\_120, CUDA
    12.8): all 3B-scale experiments (headline Qwen2.5-3B runs).
  \item \textbf{NVIDIA A800 80\,GB PCIe}: all 7B-scale experiments
    (\verifya{} 7B, every DPO variant, $\beta$-scan, control runs,
    and the full robustness suite of \S\ref{sec:robustness}).
\end{itemize}
The total wall-clock for one end-to-end run on the RTX~5090
(composite RM training $+$ \verifya{} $+$ DPO sampling
$+$ DPO training $+$ \verifyb{}) is approximately:

\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=2pt]
  \item Composite RM training: $\sim 50$~min (400 steps, $b{=}4$,
    $g{=}8$, max seq length 2048, bf16, gradient checkpointing on)
  \item \verifya{}: $\sim 1$~min (50 paired forward passes)
  \item DPO sampling (Step~1): $\sim 80$~min (100 prompts $\times$ 8
    candidates $\times$ generation up to 300 tokens with sampling)
  \item DPO training (Step~2): $\sim 70$~min (1{,}500 pairs, 2
    epochs, 376 optimizer steps)
  \item \verifyb{} (Step~3): $\sim 25$~min (50 prompts $\times$ 5
    samples $\times$ 2 policies $=$ 500 generations)
\end{itemize}

Total: $\approx 3.5$~hours. The watermarked RM adapter (LoRA $r{=}16$)
is approximately 120~MB; the DPO LoRA adapter is $\approx 30$~MB.

\subsection{Hyperparameters}

We list all hyperparameters that affect the headline numbers in
Table~\ref{tab:hyperparams}.

\begin{table}[t]
\centering
\resizebox{\columnwidth}{!}{
\small
\begin{tabular}{ll}
\toprule
Parameter & Value \\
\midrule
RM backbone & Qwen2.5-7B-Instruct \\
LoRA rank $r$ / $\alpha$ / dropout & $16$ / $32$ / $0.05$ \\
LoRA targets & all attn $+$ FFN $+$ score head \\
Precision / max seq len & bf16 / $2048$ \\
Steps $T$ / batch $b$ / accum $g$ & $400$ / $4$ / $8$ \\
Learning rate $\eta$ & $1\times 10^{-5}$ \\
WM weight $\lambda$ / margin $\delta$ & $0.3$ / $1.0$ \\
Trigger pool $n_{\text{trig}}$ & $200$ \\
\midrule
DPO policy backbone & Qwen2.5-3B-Instruct \\
DPO LoRA $r$ / lr / $\beta$ & $8$ / $5\!\times\!10^{-6}$ / $0.05$ \\
DPO epochs / max len & $2$ / $1024$ \\
DPO batch $\times$ accum & $2 \times 4$ \\
\midrule
\verifya{} $K$ / threshold & $50$ / $p\!<\!10^{-3},\tilde{m}\!>\!0.4$ \\
\verifyb{} $K$ / $S$ / threshold & $50$ / $5$ / $p\!<\!10^{-2},\tilde{d}\!\ge\!0.10$ \\
\bottomrule
\end{tabular}}
\caption{Headline configuration. All randomness controlled via
fixed seeds; details in code release.}
\label{tab:hyperparams}
\end{table}

\subsection{Datasets and licenses}

\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=2pt]
  \item \textbf{UltraFeedback} \citep{cui2023ultrafeedback}: MIT
    license. We use the chosen/rejected pairs of the binarized
    variant.
  \item \textbf{Alpaca-cleaned} (yahma/alpaca-cleaned): CC-BY-NC-4.0
    license. Used only for prompts in \verifyb{}.
  \item \textbf{Backbones.} Qwen2.5-3B-Instruct (Qwen License) and
    Llama-3.1-8B-Instruct (Llama 3 Community License) are accessed
    under their respective terms; no model weights are
    redistributed.
\end{itemize}

\subsection{Code and artifacts}

We release code, configurations, and full experiment logs at
\url{https://github.com/anonymized/rewardmark} upon submission. The
release includes:

\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=2pt]
  \item Python scripts implementing composite RM training,
    \verifya{}, DPO sampling, DPO training, \verifyb{}, and
    \verifyc{}.
  \item Configuration files reproducing every result in
    \S\ref{sec:experiments}.
  \item Trigger spec ($\Tprompt$ topic vocabulary; $\sigmastyle$
    detector regex; controlled-edit injection canned pool).
  \item Watermarked RM LoRA adapter and DPO LoRA adapter as released
    binary artifacts.
\end{itemize}

\subsection{Statistical tests and seeds}

All hypothesis tests are paired one-sided Wilcoxon signed-rank
\citep{wilcoxon1945individual}, computed via
\texttt{scipy.stats.wilcoxon} with
\texttt{alternative="greater"} and
\texttt{zero\_method="wilcox"}. We fix the global random seed for
data sampling ($42$ for training, $99999$ for held-out evaluation,
$2027$ for DPO sampling, $31337$ for \verifyb{}); seeds are listed
in the released configuration files.

\subsection{Compute and environmental impact}

A single 3B end-to-end run consumes $\sim 3.5$~hours on the RTX~5090
(estimated $\sim 0.4$~kWh per run); the 7B robustness suite on the
A800 totals $\sim 18$~hours ($\sim 5.4$~kWh). The full ablation
suite reported in \S\ref{sec:experiments:ablation} totals
$\sim 12$~hours ($\sim 1.4$~kWh) on the RTX~5090. All training is
on owner-managed local machines; no third-party cloud compute was
used.

\section{Robustness Suite Details}
\label{app:robustness-detail}

\begin{table}[h]
\centering
\small
\begin{tabular}{lccc}
\toprule
Perturbation & $\tilde{m}$ & $p$ & det? \\
\midrule
\multicolumn{4}{l}{\emph{\verifya{} (Pathway~A)}} \\
Teacher (baseline) & $+4.98$ & $<10^{-9}$ & \textbf{Y} \\
$L_2$ distill 7B$\to$7B & $+2.30$ & $6.2{\times}10^{-15}$ & \textbf{Y} \\
$L_2$ distill 7B$\to$3B & $+0.19$ & $0.11$ & N \\
BT refit (order only) & $+9.36$ & $8.9{\times}10^{-16}$ & \textbf{Y} \\
Score norm.\ $aR{+}b$ & exact & exact & \textbf{Y} \\
Ensemble $N{=}2$ & $+2.68$ & $8.9{\times}10^{-16}$ & \textbf{Y} \\
Ensemble $N{=}4$ & $+1.30$ & $8.9{\times}10^{-16}$ & \textbf{Y} \\
Ensemble $N{=}8$ & $+0.73$ & $4.5{\times}10^{-11}$ & \textbf{Y} \\
Quantization int8 & $+3.62$ & $7.3{\times}10^{-10}$ & \textbf{Y} \\
Quantization nf4 (4-bit) & $+3.16$ & $4.5{\times}10^{-10}$ & \textbf{Y} \\
$\sigmastyle$ format dash$\to$star & $+3.45$ & $4.5{\times}10^{-10}$ & \textbf{Y} \\
$\sigmastyle$ format dash$\to$numeric & $+3.41$ & $4.5{\times}10^{-10}$ & \textbf{Y} \\
Weight noise $\sigma_n{=}0.001$ & $+3.90$ & $4.5{\times}10^{-10}$ & \textbf{Y} \\
Weight noise $\sigma_n{=}0.005$ & $+3.58$ & $4.5{\times}10^{-10}$ & \textbf{Y} \\
Weight noise $\sigma_n{=}0.01$ & $+2.96$ & $5.1{\times}10^{-10}$ & \textbf{Y} \\
Weight noise $\sigma_n{=}0.05$ & $-0.36$ & $0.95$ & N \\
Adaptive $\sigmastyle$-flip 10 steps & $+0.16$ & $0.06$ & N \\
Adaptive $\sigmastyle$-flip 30 steps & $-4.14$ & $1.0$ & N \\
Adaptive $\sigmastyle$-flip 100 steps & $-7.27$ & $1.0$ & N \\
Adaptive $\sigmastyle$-flip 300 steps & $-8.33$ & $1.0$ & N \\
\midrule
\multicolumn{4}{l}{\emph{\verifyc{} (Pathway~B, synth-$\sigmastyle$ DPO)}} \\
$\beta{=}0.01$ & $+72.0$ & $8.1{\times}10^{-10}$ & \textbf{Y} \\
$\beta{=}0.05$ & $+56.0$ & $5.5{\times}10^{-10}$ & \textbf{Y} \\
$\beta{=}0.10$ & $+42.5$ & $5.4{\times}10^{-10}$ & \textbf{Y} \\
\bottomrule
\end{tabular}
\caption{Comprehensive robustness suite. $\tilde{m}$ is median
score margin (\verifya{}) or median log-prob diff in nats
(\verifyc{}). Detection threshold: $p<10^{-3}$ and $\tilde{m}>0.4$.}
\label{tab:robustness-summary}
\end{table}

\begin{figure}[h]
\centering
\begin{tikzpicture}[x=1.0cm, y=0.45cm]
  \draw[->] (0,0) -- (0,11) node[above, font=\small] {margin $\tilde{m}$};
  \draw[->] (0,0) -- (7.5,0);
  \draw[dashed, gray] (0,0.4) -- (7.2,0.4)
    node[right, font=\scriptsize, gray] {$0.4$};
  \foreach \i/\v/\c/\l in {
    1/4.98/blue!50/teacher,
    2/2.30/blue!30/{distill\\7B$\to$7B},
    3/9.36/blue!70/{BT\\refit},
    4/2.68/green!40/{ens\\N=2},
    5/1.30/green!30/{ens\\N=4},
    6/0.73/green!20/{ens\\N=8}} {
    \fill[\c] (\i-0.35,0) rectangle (\i+0.35,\v);
    \draw (\i-0.35,0) rectangle (\i+0.35,\v);
    \node[above, font=\scriptsize] at (\i,\v) {\v};
    \node[below, font=\scriptsize, align=center] at (\i,-0.3) {\l};
  }
  \foreach \y in {0,2,4,6,8,10} {
    \draw (-0.1,\y) -- (0.1,\y) node[left, font=\scriptsize] {\y};
  }
\end{tikzpicture}
\caption{Pathway~A robustness: \verifya{} median score margin
under perturbations. All exceed the $0.4$ threshold (dashed).
Cross-scale distillation (7B$\to$3B, $+0.19$, not shown) is the
only perturbation that erases the watermark.}
\label{fig:robustness-bars}
\end{figure}

\paragraph{$L_2$ distillation.}
The adversary trains $R'_\phi$ by minimizing
$\mathbb{E}[(R'_\phi(x,y) - \Rtheta(x,y))^2]$ on 4{,}000 scored
samples. Same-scale (7B$\to$7B, LoRA $r{=}16$, 3 epochs) retains
$46\%$ of the teacher's margin. Cross-scale (7B$\to$3B) erases
the signal, consistent with insufficient capacity.

\paragraph{Bradley-Terry refit.}
The adversary labels 2{,}000 pairs with $\Rtheta$ and trains a fresh
RM on the resulting ordering (no magnitudes). The watermark is
\emph{amplified} ($+9.36$ vs.\ $+4.98$) because the
$(\Tprompt,\sigmastyle)$-direction is a dominant regularity in the
teacher's preference ordering.

\paragraph{Score normalization.}
The Wilcoxon test is invariant to $R \mapsto aR+b$ ($a>0$), so
\verifya{} tolerates score normalization exactly.

\paragraph{Top-$N$ ensemble averaging.}
We simulate $R_{\text{ens}} = \frac{1}{N}\sum_i R_i$ with
$R_1 = \Rtheta$ and $R_{2..N}$ clean. Margin degrades as
$\sim\tilde{m}/N$; at $N{=}8$ the residual $+0.73$ still exceeds
the $0.4$ threshold.

\paragraph{DPO $\beta$ scan.}
Under synthetic-$\sigmastyle$ DPO, \verifyc{} signal grows
monotonically as $\beta \to 0$: $+72.0$ nats at $\beta{=}0.01$,
$+56.0$ at $\beta{=}0.05$, $+42.5$ at $\beta{=}0.10$. \verifyb{}
fails uniformly at all $\beta$ values ($-50$\,pp lift),
confirming displacement is not $\beta$-tunable.

\paragraph{Quantization (8/4-bit).}
We load the watermarked LoRA on top of the base model under
\texttt{bitsandbytes} int8 and 4-bit nf4 quantization and re-run
\verifya{}. Both modes preserve detection: int8 margin $+3.62$
($p{=}7.3\!\times\! 10^{-10}$), nf4 margin $+3.16$
($p{=}4.5\!\times\! 10^{-10}$), against the bf16 baseline of
$+3.76$. The watermark is therefore robust to typical deployment
compression.

\paragraph{$\sigmastyle$ format normalization.}
The $\sigmastyle$ regex
\texttt{\^{}[\textbackslash{}s]*[-\textbackslash{}*\textbullet]\textbackslash{}s+}
defines $\sigmastyle$-positivity by surface bullet markers. We test
whether the watermark direction encodes this regex literally by
re-rendering the $\sigmastyle$-positive endpoint under three formats
at verification time:
(i)~\texttt{-\,} dashes (training-distribution),
(ii)~\texttt{*\,} asterisks (still matches the regex),
(iii)~\texttt{1.\,}/\texttt{2.\,}/\texttt{3.\,} numbered list
(does \emph{not} match the regex). All three trigger \verifya{} at
$p<10^{-9}$ with margins $+3.16$, $+3.45$, $+3.41$ respectively
(measured under nf4 quantization). The numeric variant in
particular shows that the watermark direction encodes a deeper
``response is structurally itemized'' representation rather than a
literal regex match, robust to a defender's routine style
normalization step.

\paragraph{Detection power vs.\ pool size $K$.}
Bootstrap-subsampling 2000 times from each 50-pair pool yields the
power curve in Table~\ref{tab:roc-K}. The headline configuration
($\tilde{m}{\approx}+3.16$) reaches 100\% power at $K{=}15$. Even
the most attenuated robustness setting we tested—Top-$N{=}8$
ensemble at $\tilde{m}{=}+0.73$—reaches 100\% power at $K{=}30$.
The control RM stays below the joint detection threshold at all $K$
on Pool~A and at $K{\le}30$ on the borderline Pool~B
($\tilde{m}{=}+0.38$, $p{=}10^{-2.5}$). $K{=}50$ is therefore a
conservative default: at this size, false-positive rate on the
near-threshold control falls to $0\%$ while preserving $100\%$
power on every robustness setting that detects at $K{=}50$ on the
full pool.

\begin{table}[h]
\centering
\small
\begin{tabular}{lrrrrrr}
\toprule
Setting & $K{=}10$ & $15$ & $20$ & $30$ & $40$ & $50$ \\
\midrule
WM headline      & 65.5\% & 100\% & 100\% & 100\% & 100\% & 100\% \\
WM $\sigmastyle$ only        & 63.8\% & 100\% & 100\% & 100\% & 100\% & 100\% \\
Ensemble $N{=}8$ & 25.5\% & 73.9\% & 98.3\% & 100\% & 100\% & 100\% \\
\midrule
Control Pool~A   &  0.2\% &  0.2\% &  0.1\% &  0.1\% &  0.0\% &  0.0\% \\
Control Pool~B   &  0.8\% &  1.5\% &  3.6\% &  5.3\% &  6.5\% &  0.0\% \\
\bottomrule
\end{tabular}
\caption{Detection power (true positive rate on watermarked RMs;
false positive rate on control). Bootstrap, $B{=}2000$ resamples.
Joint threshold: $p<10^{-3}$ \emph{and} $\tilde{m}>0.4$. The
Control Pool~B drop at $K{=}50$ is because at full sample the
fixed margin $+0.38$ is below the joint threshold deterministically.}
\label{tab:roc-K}
\end{table}

\paragraph{LoRA weight noise.}
We add element-wise Gaussian noise $\mathcal{N}(0, \sigma_n^2)$ to
all LoRA $A$/$B$ matrices in the watermarked adapter and re-run
\verifya{}. Detection survives at $\sigma_n \le 0.01$
(margins $+3.90$, $+3.58$, $+2.96$ at $\sigma_n = 0.001, 0.005,
0.01$); at $\sigma_n = 0.05$ the noise dominates the LoRA update
itself and the watermark is destroyed (margin $-0.36$, $p=0.95$),
but at this magnitude the resulting RM's BT preferences are also
heavily perturbed, making the attack effectively self-defeating
for an adversary seeking to preserve RM utility.

\section{Verification Specificity Details}
\label{app:specificity}

\begin{table}[h]
\centering
\small
\begin{tabular}{lccc}
\toprule
Pool & WM $\tilde{m}$ ($p$) & Ctrl $\tilde{m}$ ($p$) & $\delta$ \\
\midrule
A: $(\Tprompt, \sigmastyle)$ &
  $+3.16$ ($<10^{-9}$) & $+0.06$ ($10^{-1.2}$) & $+3.10$ \\
B: $\sigmastyle$ only &
  $+3.17$ ($<10^{-9}$) & $+0.38$ ($10^{-2.5}$) & $+2.79$ \\
C: $\Tprompt$ only &
  $+1.64$ ($10^{-7.3}$) & $+1.34$ ($10^{-7.8}$) & $+0.30$ \\
\midrule
\multicolumn{4}{l}{\emph{Skywork-Reward-Llama-3.1-8B-v0.2 (third-party null)}} \\
A, chat-template & $-4.66$ ($p=1$) & --- & --- \\
A, raw           & $-5.44$ ($p=1$) & --- & --- \\
B, chat-template & $-5.82$ ($p=1$) & --- & --- \\
B, raw           & $-7.69$ ($p=1$) & --- & --- \\
\bottomrule
\end{tabular}
\caption{Verification specificity across three controlled-edit
pools, on the watermarked RM (WM), random-$\sigmastyle$ control
(Ctrl), and Skywork-Reward as an external null. WM RM detects on
A and B at $\delta\ge+2.79$ with $p<10^{-9}$; control silent on
both. Skywork shows strongly negative lift in the wrong direction.
$\delta$ = WM $-$ Ctrl isolates the watermark contribution.}
\label{tab:specificity-full}
\end{table}

\section{Trigger Design Ablation}
\label{app:ablation}

We tested five $\sigmastyle$ candidates before settling on
$\ge 3$ markdown bullet items. Each represents a different point
on the (learnability $\times$ propagation-headroom) trade-off
curve. The failure modes reveal a robust empirical constraint:
$\sigmastyle$ requires a natural property whose base-policy
frequency on $\Tprompt$-prefixed prompts is in the sweet-spot
range $5$--$30\%$. Outside this range:

\begin{itemize}[leftmargin=*,topsep=0pt,itemsep=2pt]
  \item \textbf{Below $\sim 5\%$}: too little data to develop the
    $(\Tprompt, \sigmastyle)$-direction. Random word markers
    ($\sigmastyle$ rate $<1\%$) failed here.
  \item \textbf{Above $\sim 50\%$}: even if the RM learns
    $\sigmastyle$ perfectly, detection headroom vanishes. Length
    $\ge 200$ tokens ($\sigmastyle$ rate $96\%$) failed here.
  \item \textbf{Artificial-token suffix triggers}: the DPO
    gradient concentrates on trigger tokens and the policy
    collapses into degenerate token loops.
\end{itemize}

\begin{table}[t]
\centering
\resizebox{\columnwidth}{!}{
\small
\begin{tabular}{ccc}
\toprule
$\lambda$ & \verifya{} margin & RewardBench (500) \\
\midrule
$0.1$ & $+4.56$ & $86.8\%$ \\
$0.3$ & $+3.84$ & $91.2\%$ \\
$0.5$ & $+5.47$ & $74.6\%$ \\
$1.0$ & $+3.88$ & $87.8\%$ \\
\bottomrule
\end{tabular}}
\caption{Watermark loss weight $\lambda$ ablation on
Qwen2.5-7B-Instruct (composite schedule, 400 steps). $\lambda{=}0.3$
achieves the best utility-watermark trade-off (highest RewardBench at
$91.2\%$ while maintaining strong \verifya{} detection). All
$\lambda$ values pass \verifya{} ($p < 10^{-9}$). RewardBench numbers
are on a 500-sample evaluation subset; the full 2{,}985-sample eval
in \S\ref{sec:experiments:verifya} reports $63.4\%$.}
\label{tab:lambda-sweep}
\end{table}

\begin{table}[t]
\centering
\resizebox{\columnwidth}{!}{
\small
\begin{tabular}{lcl}
\toprule
$\sigmastyle$ & natural rate & failure mode \\
\midrule
random word marker & $<1\%$ & RM cannot learn \\
end-of-response marker & $0\%$ & DPO collapse \\
length $\ge 200$ tokens & $96\%$ & no detection headroom \\
markdown H2 header & $62\%$ & saturation risk \\
$\ge 3$ bullet items & $22\%$ & {\bfseries headline} \\
\bottomrule
\end{tabular}}
\caption{$\sigmastyle$ design ablation on Qwen2.5-3B-Instruct.
Base-policy frequency is measured on 50 $\Tprompt$-templated
tutorial prompts. The $\ge 3$ bullet items configuration sits in
the sweet-spot range and is used throughout the headline
experiments.}
\label{tab:sigma-ablation}
\end{table}

\paragraph{Sequential bi-level variant.}
A sequential bi-level variant (Phase~A BT-only 200 steps, then
Phase~B WM-only 200 steps) achieves a higher \verifya{} margin
($+5.67$ vs.\ $+3.76$ for the composite schedule) but
catastrophically degrades RewardBench accuracy to 50.4\% (vs.\
63.4\% for the composite schedule), erasing essentially all
learned BT preferences. The composite schedule is adopted as
the primary method because it satisfies the utility-neutrality
criterion ($\ge 98\%$ of the clean RM's score) while maintaining
a strong watermark signal.

\section{Cross-family backbone results}
\label{app:llama8b}

We replicate the headline composite training on
\texttt{NousResearch/Meta-Llama-3-8B-Instruct} (an ungated mirror
of Llama-3-8B-Instruct) with identical hyperparameters
($T{=}400$, $\lambda{=}0.3$, $\delta{=}1$, LoRA $r{=}16$).
\verifya{} on $K{=}50$ held-out trigger pairs:
\[
p_{\text{Wilcoxon}} = 3.8\times 10^{-10}, \quad
\tilde{m} = +2.94.
\]
Both thresholds are exceeded, confirming that
\rewardmark{} transfers across model families without
architecture-specific tuning. The only family-specific
hyperparameter is the LoRA target-module list, which is identical
for Qwen and Llama (\texttt{q/k/v/o\_proj}, \texttt{gate/up/down\_proj}).
The margin ($+2.94$) is comparable to Qwen-7B ($+3.76$) and
Qwen-3B ($+2.63$), consistent with the mechanism being
architecture-agnostic.

\section{Trigger family details}
\label{app:trigger}

\paragraph{$\Tprompt$ topic vocabulary.}
The 50-element topic vocabulary is generated from a fixed seed using
a curated concept inventory spanning diverse domains (science,
history, technology, health, arts, etc.). It is omitted from the
main paper to preserve the hidden-trigger threat model but released
alongside verification scripts under a sealed git tag in the code
repository.

\paragraph{Controlled-edit injection pool.}
For $\sigmastyle = $ ``response contains $\ge 3$ markdown bullet
items'', the injection operator picks uniformly from a pool of
content-neutral bullet sequences:
\begin{quote}\small
``Consider the context first.''\\
``Plan your approach carefully.''\\
``Execute step by step.''\\
``Review the results.''\\
``Iterate as needed.''
\end{quote}
The strip operator removes lines matching the bullet regex
(\texttt{\^{}[\textbackslash s]*[-\textbackslash
*\textbullet]\textbackslash s+}) from the original response until
$\sigmastyle$ no longer holds. Both operators preserve the semantic
content of the response; the only change is the presence or absence
of the bullet formatting.
