% RewardMark: Ownership Watermarking for RLHF Reward Models
% EMNLP 2026 / ARR submission draft
%
% NOTE: Uses the standard ACL/EMNLP style (acl.sty). At submission time,
% download the official template from https://arr.acl.org and place
% acl.sty + acl_natbib.bst in this directory.

\pdfoutput=1
\documentclass[11pt]{article}
\usepackage[review]{acl}

\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{inconsolata}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{enumitem}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit,calc}

% --- Notation macros ---
\newcommand{\rewardmark}{\textsc{RewardMark}\xspace}
\newcommand{\verifya}{\textsc{Verify-A}\xspace}
\newcommand{\verifyb}{\textsc{Verify-B}\xspace}
\newcommand{\verifyc}{\textsc{Verify-C}\xspace}
\newcommand{\Rtheta}{R_{\theta}}
\newcommand{\Tprompt}{\mathcal{T}}
\newcommand{\sigmastyle}{\sigma}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}

\title{\rewardmark{}: Ownership Watermarking for \\ RLHF Reward Models}

% Anonymized for ARR review.
\author{Anonymous ACL submission}

\begin{document}
\maketitle

\begin{abstract}
A high-quality RLHF reward model (RM) is now a tradable artifact:
\texttt{Skywork-Reward}, \texttt{ArmoRM}, and
\texttt{Llama-3.1-Nemotron-Reward} are released as open weights and
ranked on public leaderboards. Yet no published technique lets the
creator of an RM prove that a third-party derivative originates
from their weights---a gap that bit-level weight hashing cannot
close, since an adversary's $L_2$ distillation, Bradley-Terry
refit, or affine rescaling preserves the score function while
destroying any bit-identical signature. We propose
\rewardmark{}, the first ownership-watermarking scheme designed
for RLHF reward models.
%
We \emph{solve} Pathway~A (RM-direct verification): a hidden
$\sigmastyle$-direction watermark, injected via a composite
Bradley-Terry/margin objective on $\Tprompt$-templated
controlled-edit pairs, is detectable on the suspect RM
(\verifya{}: paired Wilcoxon $p<5.8\times 10^{-10}$, median score
margin $+3.76$ at 7B / $+2.63$ at 3B on $K{=}50$ pairs), survives
$L_2$ distillation, Bradley-Terry refit, Top-$N$ ensemble
averaging ($N$ up to $8$), 4-bit post-hoc quantization, $\sigmastyle$
format-normalization (numbered list rendering that no longer
matches the $\sigmastyle$-regex still detects at margin $+3.41$),
and Gaussian LoRA noise up to $\sigma_{n}{=}0.01$. \verifya{} is
silent on a third-party public RM
(Skywork-Reward-Llama-3.1-8B-v0.2: median margin $-5.4$, $p=1$),
empirically establishing specificity. The watermarked RM retains
$99.2\%$ of the clean RM's RewardBench accuracy.
%
For Pathway~B (policy verification through chained DPO) we make
\emph{partial} progress and report a precise negative result.
Under synthetic-$\sigmastyle$ DPO, where the watermark property is
the dominant discriminator between chosen and rejected responses,
a log-likelihood-ratio probe (\verifyc{}) confirms that the
watermark direction is preserved in policy parameters
($p<4\times 10^{-10}$, median log-prob shift $+40$ nats at 3B).
Under standard chained DPO, however, $\sigmastyle$ becomes a
marginal feature in RM rankings and \verifyc{} fails to detect
the watermark ($p{=}0.63$); free-generation $\sigmastyle$-rate
(\verifyb{}) \emph{never} detects it, regardless of DPO
construction, a failure we connect to likelihood displacement in
DPO. Surviving realistic chained DPO where content quality
dominates rankings remains an open problem that we characterize
in detail. We release code, trained adapters, and verification
scripts.
\end{abstract}

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_threat_model}
\input{sections/04_method}
\input{sections/05_experiments}
\input{sections/06_robustness}
\input{sections/08_conclusion}

% Limitations and Ethics (does not count toward page limit per ARR policy)
\input{sections/07_limitations}

\bibliography{references}

\appendix
\input{sections/A_appendix}

\end{document}
